% \font\x1 = cmbx10 scaled \magstephalf
% \font\x2 = cmbxti10 scaled \magstephalf 

\leftline{Example 1 : \quad Simple text}
\vglue .5\baselineskip          %skip an extra half line
``Then you can't call him Winnie?''

``I don't.'' \footnote \dag{``He's Winnie-ther-Pooh. Don't you know what '{\it ther}' means?''}

``Ah,~yes, now I do.'' I said quickly; and I---hope you do--too, because it is all the explanation you are going to get.

{\parindent = 0pt \parskip = 6pt
Sometimes Winnie-the-Pooh likes a game of some sort of when he comes
down stairs, and sometimes he likes to sit quietly in front of the fire and
listen to a story.

\leftline{This evening ---}
\par
}

``What about a story?'' said Christopher Robin.

``What about a story?'' I said.

``Could you very sweetly tell Winnie-the-Pooh one?''

{\narrower ``I suppose I could,'' I said. ``What sort of stories does he
likes?''

``About himself. Because he's {\it that} sort of Bear''

``Oh, I see.''}

\vskip 1pc
``So could you very sweetly?''

``I'll try,'' I said. So I tried.
{\narrower \narrower Once upon a time, a very long time ago now, about
last Friday, Winnie-the-Pooh lived in a forest all by himself under the
name of Sanders.}

{\it ('What does ``under the name'' mean?' asked Christopher Robin.)

'It means he had the name over the door in gold letters and lived under it.'
}

{\leftskip .5in {\it ``Winnie-the-Pooh wasn't quite sure'', said Christopher Robin.}\par}
{\rightskip .5in {\it ``Now I am,'' said a growly voice.
``Then I will go on,'' said I.
\par}
One day when he was out of walking, he came to an open place in the middle of the forest,
and in the middle of this place was a large oak tree, and, from the top of the tree, there came
a loud buzzing-noise.
}


% special char
% \chardef \\ = '\\ %let \\ denote a backslash

{\it Winnie-the-Pooh sat down at the\/ {\bf foot of the tree}}, put {\bf his
head between his paws, and } begin to \/{\rm think.}

First of all he said {\sevenrm to himself: } That buzzing-noise means something.
You don't get a {\fiverm buzzing-noise like that, just buzzing and buzzing, without
meaning something. }

\medskip
\centerline{If \$ \quad there's \& \quad  a \# \quad buzzing-sound \% \quad, somebody \^ \quad
}
\centerline{
is making a buzzing-noise \~ \quad,  and $\{$ \quad the $\}$ \quad only $\backslash$
}
\centerline{
reason for that I know of is because you're a bee.
}
\medskip

\noindent Then he thought another long time, and said: ``And the only reason for
being a bee that I know of is making h\^oney.''

And then he got up, and said : ``And the only reason for making honey is so
{\it a\ss \/ I \/} can e\'at it ``. S$\Omega$ he beg$\alpha$n to climb the tree.

He climbed and he climbed and he climbed, $\sharp$ and as he climbed $\flat$ he
sang a little song $\spadesuit$ to himself $\diamondsuit$. 


% Interline spacing
\baselineskip = 2\baselineskip %doubling spacing
\parskip = \baselineskip       %skip a line between paragraph
\parindent = 3em               %Increase indentation of paragraph

% A macro that makes a nice fraction
\def\frac #1/#2{
\leavevmode
\kern.1em \raise .5ex \hbox{\the \scriptfont0 #1}
  \kern-.1em $/$
  \kern-.15em \lower .25ex \hbox{\the \scriptfont0 #2}
}

Hi \frac 1/2.
It went like this:

Isn't it funny?

How a bear likes honey?

Buzz ! Buzz! Buzz!

I wonder why he does?
\baselineskip = 0.5\baselineskip %doubling spacing
\parskip = \baselineskip       %skip a line between paragraph

% Odds and ends


\footline{\hfil {\tenit - \folio -}\hfil }
% a footer line.

Then he climbed a little further...and a little further.. and then just a little
further.

${\it 5}$-[p-(Flouro\-sul\-fonyl)ben\-zoyl]-l,
$N^6$-ethe\-no\-adeno\-sine.

By that time he had thought of another song. \medskip

{\raggedright It's a very
funny thought that, if Bears were Bees, They'd build their own nests at the {\it bottom\/ } of trees.
And that being so (if the Bees were Bears),
We shouldn't have to climb up all these stairs.
\par}

\proclaim Assertion 27. He was getting rather tired by this time, so that is
whay he sang a Complaining Song. He was nearly there now, and if he just stood
on that branch...

{\it Crack!}

``Oh, help!'' said Pooh, as he dropped ten feet to the branch below him.

\smallskip
\item{1.} ``If only I hadn't ---'' he said , as he bounced twenty feet on to the next branch.
\item{2.} ``You see, what I {\it meant} to do,`` he explained, as he turned head-over-heels, and crashed on to another branch
thirty feet below, ``what I {\it meant} to do ---''

\item{} \indent ``Of course, it {\it was} rather --'' he admitted, as he slithered very quickly through the next six branches.
\itemitem {(a)} ``It all comes, I suppose,'' he decided,
\itemitem {(b)} as he said good-bye to the last branch, spun round three times, and flew gracefully into a gorse-bush,
\item {$\bullet$} ``It all comes of liking honey so much. Oh help!''
\smallskip
\leftline{He crawled out of the gorse-bush, brusched the prickles from his nose, $\Leftarrow$}
\rightline{$\Rightarrow$ and the first person he thought of was Christopher Robin.}
\centerline{$\Rightarrow$(``Was that me?'' said Christopher Robin)$\Leftarrow$}

in an awed voice.
\bye
