\message{ !name(m.tex)}
\message{ !name(m.tex) !offset(-2) }
\documentclass[dvipsnames]{article}
% \documentclass[dvipsnames]{ctexart}

\title{Openstack}
\input{h}
% --------------------------------------------------
\begin{document}
\maketitle

\section{Consideration}

\subsection{The logical architecture}

\emoji{turtle} : OpenStack uses a microservice architecture. So it's composed of
many many services. 

About this services:

\begin{itemize}
\item Most of them of them are written in \cola{Python}.
\item All of them provide a \cola{REST API}.
\item Each service may be implemented as different components. The components
  communicate with each other using a \cola{message queue}.
\end{itemize}

\subsubsection{Keystone - identity management}

\emoji{turtle} : Keystone is the identity management service. It's the simplest
service in OpenStack.

It provides a REST API for authentication and authorization. It also provides a
service catalog, which is a list of all the services.

\emoji{parrot} : Why is it keeping it ?

\emoji{turtle} : I think probably because it needs to know which services are
accessible for a user anyway.

\subsubsection{Swift - object storage}

\emoji{turtle} : Swift is the object storage service (an object store). To store
an object, it splits the object into chunks and stores them in different
containers redundantly.

Some of the benefits of this design are:

\begin{itemize}
\item It has no central brain, and indicates no \Cola{Single Point of Failure (SPOF)}.
\item It has \cola{auto-revovery}.
\item It has inexpensive hardware that can be used for redundant storage clusters.
\end{itemize}

\emoji{parrot} : Wow, \cola{auto-recovey} is cool.

\subsubsection{Cinder - block storage}

\emoji{parrot} : What if I just want a block of bytes ? For example, for the
volume of a VM.

\emoji{turtle} : Then you can use Cinder. It provides raw volumes that can be
used as hand disks in VMs.

Some of the features of Cinder are:

\begin{itemize}
\item You can create snapshots of volumes.
\item You can create backups of volumes.
\end{itemize}

Also, like Keystone, different storage backends can be plugged into Cinder such
as from IBM, NetApp, Nexenta, and VMware.

\emoji{parrot} : What are these ?

\emoji{turtle} : I don't know. I think they are companies that provide
storage....

\begin{tcolorbox}
  \emoji{turtle} : Cinder replaces the old \Cola{Nova Volume service}. 
\end{tcolorbox}

\subsubsection{Manila - file share (since \cola{Juno})}

\emoji{turtle} : Manila is the file share service. It provides a remote file
system. In operation, it resembles the \Cola{Network File System (NFS)} or
\Cola{SAMBA} storage service that you can find in Linux.

\emoji{parrot} : Oh, I know SAMBA. I remember it's how linux can access windows
file.

\emoji{turtle} : Yes, it's a file sharing protocol.

In contrast to Cinder, it resembles the \Cola{Storage Area Network (SAN)}
service. In fact, NFS and SAMBA or the \Cola{Common Internet File System (CIFS)}
are supported as backend drivers. Manila provides the orchestration of shares on
the share services.

\emoji{parrot} : What do you mean by \cola{orchestration of shares} ?

\emoji{turtle} : I think it means that it will take care of the shares when they
are read and written by different services.

\emoji{parrot} : Oh, so what's the difference between Swift, Cinder and Manila ?

\emoji{turtle} : Let's see a table:

\begin{table}[htbp]
  \centering
  \begin{tabularx}{0.8\textwidth}{XXXX}
    \textbf{Specification} & \textbf{Swift} & Cinder & Manila \\
    \hline
    Access mode & REST API & Block device & File system \\
    Multi-access & Yes & No & Yes \\
    Persistent & Yes & Yes & Yes \\
    Accessibility & Anywhere & Within single VM & Within multiple VMs \\
  \end{tabularx}
\end{table}

\end{document}


% Local Variables:
% TeX-engine: luatex
% TeX-command-extra-options: "-shell-escape"
% TeX-master: "m.tex"
% End:
\message{ !name(m.tex) !offset(-129) }
