\documentclass[dvipsnames]{article}
% \documentclass[dvipsnames]{ctexart}


\title{Ansible}
\input{h}
\input{h-says}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,      
  urlcolor=cyan,
  pdftitle={Linux Network},
  pdfpagemode=FullScreen,
}
\urlstyle{same}

% --------------------------------------------------
\begin{document}

\maketitle

% 🦜 : It looks like this two must be after the \maketitle
\newcolumntype{s}{>{\hsize=.2\hsize}X}
\newcolumntype{c}{>{\hsize=.2\hsize \centering\arraybackslash}X}

\section{Concepts}

\begin{enumerate}
\item \cola{Playbooks}: \colz{a list of \cola{Plays}.}
  \begin{enumerate}
  \item \cola{Plays}: \colz{They contain \cola{an ordered list of Tasks,Vriables
        and Roles}. Play maps hosts to tasks.
    }
  \item \cola{Roles}: \colz{A limited distribution of reusable Ansible
      code.(tasks,handlers, variables, plugins, templates and files) for use
      inside of a Play.

      To use any Role resource, the Role itself must be imported into the Play.
    }
  \item \cola{Tasks}: \colz{The definition of an ``action'' to be applied to the
      managed host. Tasks must always be contained in a Play, directly or
      indirectly (Role, or imported task list file). You can execute a single task
      once with an ad hoc command using \texttt{ansible} or
      \texttt{ansible-console} (both create a virtual Play).
    }
  \item \cola{Handlers}: \colz{A special form of a Task, that only executes when
      notified by a previous task which resulted in a ``changed'' status.}
  \end{enumerate}
\item Modules: \colz{The code that Ansible copied to and executes on each
    managed code (when needed) to accomplish the action defined in each Task.
    Each module has a particular use, such as:
    \begin{itemize}
    \item administering users on a specific type of database 
    \item managing VLAN interfaces on a specific type of network device.
    \end{itemize}

    You can invoke a single module with a task, or invoke several different
    modules in a playbook. Ansible modules are grouped in \cola{collections}.
    For example, \texttt{command} is a module in the \texttt{ansible.builtin}
    collection. So you can do something like 
    }
\end{enumerate}

\end{document}

% Local Variables:
% TeX-engine: luatex
% TeX-command-extra-options: "-shell-escape"
% TeX-master: "m.tex"
% TeX-parse-self: t
% TeX-auto-save: t
% End: