
\section{DNS Update}
\label{sec:dns-update}

\subsection{Definitions}
\label{sec:dns-update-defs}

\begin{itemize}
\item Sub \colz{
    an \cola{authoritative server} that uses \colZt{AXFR} or \colZt{IXFR} to retrieve
    the zone and is named in the zone's NS records.
  } 
\item Main  \colz{
    an \cola{authoritative server} that is the source of \texttt{AXFR/IXFR} data for
    one or more sub.
  }
\item Primary Main \colz{
    the main server at the root of the \texttt{AXFR/IXFR} dependency graph. The
    primary master is named in the zone's \texttt{SOA MNAME} field.
    \cola{There's only one primary main per zone.}
  }
\end{itemize}

\colz{
  A domain name identifies a node within the \cola{domain name space tree
    structure}. Each node has a set (possibly empty) of Resource Records (RRs).
  All RRs having the same \texttt{NAME, CLASS, TYPE} are called a \Cola{Resource
  Record Set (RRset).}
}

\subsection{New Assigned Numbers}
\label{sec:new-assigned-num}

\cSay{
  \colz{
    Remember the \texttt{CLASS} field in the DNS header? It's extensible. And here
    we define a new one: 
  } \cola{\texttt{CLASS=NONE (254)}}.

  Also some new \texttt{RCODE}:
  \begin{align*}
    \mathtt{YXDOMAIN} &= 6\\
    \mathtt{YXRRSET} &= 7\\
    \mathtt{NXRRSET} &= 8\\
    \mathtt{NOTAUTH} &= 9\\
    \mathtt{NOTZONE} &= 10
  \end{align*}

  And a new \cola{\texttt{OPCODE}: \texttt{UPDATE (5)}}.
}

\subsection{quiz for RFC 2136}

\begin{enumerate}
\item When two RRs are considered equal? \ans{}
\item Wildcard (\texttt{'*'}) is disabled. \tf{}
\item All \texttt{UPDATE} operations must use canonical names. \tf{}
\item Three types of RRs are not allowed to be appended to an RRset, which
  three?  \ans{}
\end{enumerate}